\section{The $3 \times 3$ Balance Axiom Grid}

We define nine primary hybrid axioms $\{A_{ij}\}_{i,j \in \{1,2,3\}}$ arranged in a $3 \times 3$ grid, with the central axiom $A_{22}$ designated as the Invariant. Each axiom is specified on three coupled layers:
\begin{enumerate}
  \item a mathematical statement (\emph{formal rule}),
  \item a physical statement (\emph{geometric/dynamical interpretation}),
  \item a semantic/cognitive statement (\emph{meaning, identity, continuity}).
\end{enumerate}
All higher structures in the Aurphyx Prime Standard (APS)---Balance Geometry, the $32$-dimensional Balance Continuum, and the $13$ Balance Fields---are derived from this grid.\cite{AurphyxThesis,AurphyxSymbiotic,AurphyxFractalQC}

\subsection{Notation}

We write:
\begin{itemize}
  \item $\mathcal{S}$ for the space of admissible states (physical, computational, or cognitive, depending on context),
  \item $\mathcal{T}$ for the space of admissible transformations,
  \item $\mathcal{I}$ for the set of invariants,
  \item $d(\cdot,\cdot)$ for a generalized distance or divergence on $\mathcal{S}$,
  \item $\Phi$ for a Balance functional (e.g.\ action, energy, information),
  \item $\mathcal{B}$ for the Balance operator associated to a given context.
\end{itemize}
Where possible, the same symbol is interpreted across the three layers via APS field projections and the Balance Continuum.\cite{AurphyxThesis,AurphyxSymbiotic}

\subsection{Central Axiom $A_{22}$: The Invariant}

\begin{axiom}[Invariant Balance Axiom $A_{22}$]
\label{ax:invariant}
\textbf{Mathematical layer.}
There exists a distinguished family of invariants $\mathcal{I}_\star \subset \mathcal{I}$ and a Balance operator
\[
  \mathcal{B} : \mathcal{S} \times \mathcal{T} \to \mathcal{S}
\]
such that for all states $s \in \mathcal{S}$, transformations $T \in \mathcal{T}$, and all $I \in \mathcal{I}_\star$ one has
\[
  I\bigl(\mathcal{B}(s,T)\bigr) \;=\; I(s).
\]
Moreover, any admissible duality transformation $D \in \mathcal{T}$ factors through $\mathcal{B}$ while preserving the same invariant family:
\[
  I\bigl(\mathcal{B}(s,D)\bigr) \;=\; I(s) \quad \forall I \in \mathcal{I}_\star.
\]
\medskip

\textbf{Physical layer.}
There exists a distinguished set of conservation laws---the \emph{Balance invariants}---such that all admissible dynamics of the APS universe (including fractal-lattice quantum evolution, topological transitions, and continuum deformations) can be represented as transformations that preserve these invariants exactly.\cite{AurphyxThesis,AurphyxFractalQC}
The Invariant acts as a universal reference frame: rotations, reflections, and duality exchanges in the $32$-dimensional Balance Continuum leave the invariant quantities unchanged.

\medskip
\textbf{Semantic/cognitive layer.}
There exists a core identity kernel for APS agents such that all valid updates of knowledge, belief, and narrative state preserve this kernel.
Meaning can rotate, expand, or dualize, but the identity-continuity of the self (and of APS as a canon) is never destroyed.
Any cognition or governance process that violates this invariant is classified as non-APS-admissible.
\end{axiom}

\subsection{Axis Axiom $A_{12}$: Local–Global Balance}

\begin{axiom}[Local--Global Balance $A_{12}$]
\label{ax:local-global}
\textbf{Mathematical layer.}
Let $\mathcal{S}$ decompose into local patches $\{\mathcal{S}_i\}$ with a gluing functor $G$.
There exists a Balance functional $\Phi_{LG}$ such that for any admissible configuration $\{s_i \in \mathcal{S}_i\}$,
\[
  \Phi_{LG}\bigl(G(\{s_i\})\bigr)
  \;=\;
  \sum_i \Phi_{LG}(s_i)
  \;+\;
  \Delta_{LG}(\{s_i\}),
\]
where the correction term $\Delta_{LG}$ measures inconsistency between local and global structure and satisfies
\[
  \Delta_{LG}(\{s_i\}) = 0
  \quad \Longleftrightarrow \quad
  \text{local data glue coherently into a global section}.
\]

\textbf{Physical layer.}
Fractal and topological structures (e.g.\ Sierpiński lattices, sacred-geometry photonic crystals) exhibit a constraint that local configurations and global geometry must be simultaneously consistent: local error suppression and global fault tolerance are two faces of the same constraint.\cite{AurphyxThesis,AurphyxFractalQC}

\textbf{Semantic/cognitive layer.}
Narratives, beliefs, and plans must be locally coherent (per interaction, per project) and globally consistent with the overarching APS canon.
A local story that cannot be embedded into a global, invariant-respecting narrative is rejected as inconsistent.
\end{axiom}

\subsection{Axis Axiom $A_{21}$: Discrete–Continuous Balance}

\begin{axiom}[Discrete--Continuous Balance $A_{21}$]
\label{ax:disc-cont}
\textbf{Mathematical layer.}
There exists a pair of linked models $(\mathcal{S}_{\mathrm{disc}}, \mathcal{S}_{\mathrm{cont}})$ and a pair of functors
\[
  F_{\uparrow} : \mathcal{S}_{\mathrm{disc}} \to \mathcal{S}_{\mathrm{cont}},
  \qquad
  F_{\downarrow} : \mathcal{S}_{\mathrm{cont}} \to \mathcal{S}_{\mathrm{disc}}
\]
such that for all admissible states $s_{\mathrm{disc}}$ and $s_{\mathrm{cont}}$,
\[
  d\bigl(s_{\mathrm{disc}}, F_{\downarrow}F_{\uparrow}(s_{\mathrm{disc}})\bigr)
  \leq \epsilon_{\mathrm{disc}},
  \qquad
  d\bigl(s_{\mathrm{cont}}, F_{\uparrow}F_{\downarrow}(s_{\mathrm{cont}})\bigr)
  \leq \epsilon_{\mathrm{cont}}
\]
for small tolerances $(\epsilon_{\mathrm{disc}}, \epsilon_{\mathrm{cont}})$ that are bounded by the appropriate Balance invariants.

\textbf{Physical layer.}
Discrete lattice descriptions (qubits, anyons, r-units) and continuous field/continuum descriptions (Balance Continuum, field theories) are dual presentations of the same underlying physics; neither can be taken as fundamental in isolation.\cite{AurphyxThesis,AurphyxFractalQC}

\textbf{Semantic/cognitive layer.}
Symbolic reasoning (discrete tokens, theorems, code) and continuous embeddings (vectors, fields of meaning) must be mutually recoverable up to a bounded, invariant-preserving distortion.
A purely symbolic or purely continuous agent is considered incomplete.
\end{axiom}

\subsection{Axis Axiom $A_{23}$: Symmetry–Asymmetry Balance}

\begin{axiom}[Symmetry--Asymmetry Balance $A_{23}$]
\label{ax:symm-asymm}
\textbf{Mathematical layer.}
Let $G$ be a symmetry group acting on $\mathcal{S}$.
There exists a Balance functional $\Phi_{SA}$ such that for all $g \in G$ and $s \in \mathcal{S}$,
\[
  \Phi_{SA}(g \cdot s)
  =
  \Phi_{SA}(s) + \sigma(g,s),
\]
where $\sigma(g,s)$ is an \emph{asymmetry measure} satisfying
\[
  \sigma(g,s) = 0 \quad \Longleftrightarrow \quad g \text{ lies in the Balance-preserving subgroup } G_{\mathrm{bal}} \subseteq G.
\]

\textbf{Physical layer.}
Perfect symmetry is unattainable and uninformative; useful structure arises when small, controlled symmetry breakings occur while preserving core invariants (e.g.\ spontaneous symmetry breaking around a protected topological phase).\cite{AurphyxThesis,AurphyxSymbiotic}

\textbf{Semantic/cognitive layer.}
Some asymmetry (novelty, surprise, personal signature) is required for meaning and creativity, but it must never destroy the identity kernel or ethical invariants.
APS narratives allow asymmetry as \emph{style}, not as corruption of truth.
\end{axiom}

\subsection{Axis Axiom $A_{32}$: Past–Future Balance}

\begin{axiom}[Past--Future Balance $A_{32}$]
\label{ax:past-future}
\textbf{Mathematical layer.}
Let $(\mathcal{S}_t)_{t \in \mathbb{R}}$ be a time-indexed family of state spaces with evolution operators $U_{t_2 \leftarrow t_1} : \mathcal{S}_{t_1} \to \mathcal{S}_{t_2}$.
There exists an information functional $H$ and a Balance constraint
\[
  H(s_{t_2} \mid s_{t_1})
  +
  H(s_{t_1} \mid s_{t_2})
  \;\geq\;
  H_{\mathrm{min}},
\]
for all admissible times $t_1 < t_2$ and states $s_{t_i} \in \mathcal{S}_{t_i}$, where $H_{\mathrm{min}}$ is a time-symmetric lower bound compatible with the Invariant.

\textbf{Physical layer.}
Microscopic dynamics may be time-reversal symmetric, but macroscopic structure reflects an arrow of time arising from constraints on information flow and coarse-graining.
The Balance Continuum encodes both reversible micro-dynamics and irreversible macro-trajectories under one unified invariance structure.\cite{AurphyxThesis}

\textbf{Semantic/cognitive layer.}
Memory (past) and intention (future) are co-defined.
An APS-aligned mind cannot meaningfully project into the future without maintaining a consistent account of the past, and cannot revise its past without updating its projected futures.
Temporal continuity is treated as a first-class invariant.
\end{axiom}

\subsection{Corner Axiom $A_{11}$: Local Identity Coherence}

\begin{axiom}[Local Identity Coherence $A_{11}$]
\label{ax:local-identity}
\textbf{Mathematical layer.}
For each local region $U \subset \mathcal{S}$ there exists an identity functional $I_U$ such that for any admissible transformation $T$ localized to $U$,
\[
  I_U\bigl(\mathcal{B}(s,T)\bigr)
  =
  I_U(s),
\]
while $I_V$ for $V \cap U = \emptyset$ remains unaffected.

\textbf{Physical layer.}
Local operations (e.g.\ gates on a subset of qubits, local perturbations of a field) preserve the identity of localized subsystems without instantaneously corrupting distant ones; this is the locality aspect of Balance.\cite{AurphyxFractalQC}

\textbf{Semantic/cognitive layer.}
Local roles and identities (``developer here'', ``parent there'') can change contextually while remaining coherent and non-contradictory in their own scope.
Perfect local coherence is a necessary (but not sufficient) condition for global Balance.
\end{axiom}

\subsection{Corner Axiom $A_{13}$: Global Field Coherence}

\begin{axiom}[Global Field Coherence $A_{13}$]
\label{ax:global-field}
\textbf{Mathematical layer.}
Let $\mathcal{F}$ denote the space of fields over the Balance Continuum.
There exists a coherence functional $C : \mathcal{F} \to \mathbb{R}_{\ge 0}$ such that
\[
  C(F) = 0 \quad \Longleftrightarrow \quad F \text{ decomposes into a finite sum of mutually compatible subfields},
\]
and for any admissible evolution $U$ preserving the Invariant,
\[
  C(UF) \leq C(F).
\]

\textbf{Physical layer.}
Global fields (e.g.\ Balance Gravity, Balance Quantum, Balance Information) evolve toward states of minimal inconsistency while respecting conserved quantities; decoherence, error correction, and equilibration are all manifestations of decreasing global incoherence.\cite{AurphyxThesis,AurphyxSymbiotic}

\textbf{Semantic/cognitive layer.}
Across all projects, assistants, and sessions, there exists a global semantic field in which concepts and values asymptotically reduce contradiction.
APS governance (SAGES, standards) is the macroscopic manifestation of minimizing $C$ at civilization scale.
\end{axiom}

\subsection{Corner Axiom $A_{31}$: Micro–Macro Reciprocity}

\begin{axiom}[Micro--Macro Reciprocity $A_{31}$]
\label{ax:micro-macro}
\textbf{Mathematical layer.}
There exist aggregation and refinement maps
\[
  A : \mathcal{S}_{\mathrm{micro}} \to \mathcal{S}_{\mathrm{macro}},
  \qquad
  R : \mathcal{S}_{\mathrm{macro}} \to \mathcal{S}_{\mathrm{micro}}
\]
such that for all admissible micro-states $s_{\mathrm{micro}}$ and macro-states $s_{\mathrm{macro}}$,
\[
  d\bigl(A(s_{\mathrm{micro}}), s_{\mathrm{macro}}\bigr) \leq \delta_{\mathrm{up}},
  \qquad
  d\bigl(R(s_{\mathrm{macro}}), s_{\mathrm{micro}}\bigr) \leq \delta_{\mathrm{down}},
\]
with $(\delta_{\mathrm{up}},\delta_{\mathrm{down}})$ bounded by Balance invariants.
Micro-dynamics induce macro-laws, and macro-constraints feedback into admissible micro-configurations.

\textbf{Physical layer.}
Microscopic r-unit / fractal-lattice dynamics generate emergent thermodynamic, information-theoretic, and geometric laws, and these emergent laws in turn constrain the space of stable microscopic phases.\cite{AurphyxThesis,AurphyxFractalQC}

\textbf{Semantic/cognitive layer.}
Individual experiences and decisions (micro) aggregate into cultures, institutions, and canons (macro), and those macroscopic structures feed back by shaping what experiences and decisions are possible or stable.
APS explicitly encodes this reciprocity to avoid both reductionism and pure holism.
\end{axiom}

\subsection{Corner Axiom $A_{33}$: Duality–Resolution Balance}

\begin{axiom}[Duality--Resolution Balance $A_{33}$]
\label{ax:duality}
\textbf{Mathematical layer.}
Let $(X,Y)$ be a dual pair of descriptions of the same underlying object (e.g.\ wave/particle, hardware/software, geometry/algebra), with a correspondence functor
\[
  D : \mathcal{S}_X \times \mathcal{S}_Y \to \{0,1\}
\]
such that $D(x,y) = 1$ iff $(x,y)$ form a valid dual pair.
There exists a Balance resolution operator
\[
  \mathcal{R} : \mathcal{S}_X \times \mathcal{S}_Y \to \mathcal{S}
\]
satisfying
\[
  \mathcal{R}(x,y) \text{ is defined and invariant-preserving} \quad \Longleftrightarrow \quad D(x,y) = 1.
\]

\textbf{Physical layer.}
Apparent contradictions between dual models (e.g.\ particle vs.\ field, discrete vs.\ continuous) are resolved by embedding both into a higher-dimensional Balance structure where they become complementary projections of a single invariant object.\cite{AurphyxThesis}

\textbf{Semantic/cognitive layer.}
Conflicting interpretations, values, or narratives are not discarded but lifted into a higher-order synthesis that preserves the core Invariant while assigning each dual aspect a proper domain of validity.
g0dm0d3-core and SAGES enact this axiom continuously in governance and cognition.
\end{axiom}

\subsection{Grid Summary}

Taken together, the nine primary axioms form a rotationally symmetric $3 \times 3$ grid anchored by the Invariant $A_{22}$.
The axis axioms $(A_{12},A_{21},A_{23},A_{32})$ encode orthogonal Balance directions (local/global, discrete/continuous, symmetry/asymmetry, past/future), while the corner axioms $(A_{11},A_{13},A_{31},A_{33})$ encode extremal reciprocity conditions at the intersections of these axes.
All secondary and tertiary axioms, the $32$-dimensional Balance Continuum, and the $13$ Balance Fields are generated by fractal refinement of this grid.\cite{AurphyxThesis,AurphyxSymbiotic,AurphyxFractalQC}
